### 2021-01-06 02_abstract/abstract_english.tex
Adapted file permissions
Added figures for ESMValTool contributions (chapter 3)

%=============================================================================%
%                                Dissertation                                 %
%                               Manuel Schlund                                %
%                                  (c) 2020                                   %
%=============================================================================%
%                              01 - Introduction                              %
%=============================================================================%

\chapter{Introduction}

Lorem ipsum dolor sit amet, consectetur adipiscing elit, sed do eiusmod tempor incididunt ut labore et dolore magna aliqua. Hendrerit gravida rutrum quisque non tellus orci ac auctor augue. Eros donec ac odio tempor orci dapibus. Ultrices neque ornare aenean euismod elementum nisi. Vivamus arcu felis bibendum ut tristique et egestas quis ipsum. Elit ut aliquam purus sit amet luctus. Et malesuada fames ac turpis egestas sed. Egestas dui id ornare arcu odio ut sem nulla pharetra. Porta lorem mollis aliquam ut porttitor leo a diam. Ultricies tristique nulla aliquet enim. Urna nunc id cursus metus aliquam. Faucibus a pellentesque sit amet porttitor. Tincidunt ornare massa eget egestas purus. Venenatis tellus in metus vulputate eu scelerisque felis imperdiet. Posuere urna nec tincidunt praesent. Risus pretium quam vulputate dignissim suspendisse in.

Aliquam id diam maecenas ultricies mi eget. Mauris pharetra et ultrices neque ornare aenean euismod. Vitae suscipit tellus mauris a diam maecenas. Massa massa ultricies mi quis hendrerit. Tincidunt eget nullam non nisi est sit amet. Vitae purus faucibus ornare suspendisse sed nisi. Enim sit amet venenatis urna. Malesuada fames ac turpis egestas. Viverra nam libero justo laoreet sit amet. Ornare massa eget egestas purus viverra accumsan in nisl nisi. Ac feugiat sed lectus vestibulum mattis.

\autocite{Schlund2020}
\autocite{Schlund2020a}

>>>>

%=============================================================================%
%                                Dissertation                                 %
%                               Manuel Schlund                                %
%                                  (c) 2020                                   %
%=============================================================================%
%                             Abstract (English)                              %
%=============================================================================%

\chapter*{Abstract (English version)}

TBA.

### 2021-01-06 ch02_scientific_background/scientific_background.tex
Added figures for ESMValTool contributions (chapter 3)

\begin{figure}[t]
  \centering
  \includegraphics[width=\FigureWidth{}]{
    ch02_scientific_background/figs/gregory_regression.pdf}
  \caption{Global and annual mean net \acf{TOA} radiation $N$ \vs{} the change
    in global and annual mean near-surface air temperature $\Delta T$ for 150
    years of a simulation with an abrupt quadrupling of the atmospheric
    \ac{CO2} concentration (\nxcotwo{4}) for the \acs{CMIP}5 multi-model mean
    (blue circles). To account for energy leakage and model drift, a linear
    fit of the corresponding pre-industrial control run is subtracted from the
    \nxcotwo{4} simulation. As given by \cref{eq:02:N_vs_delta_t}, the slope
    of the linear regression (black line) corresponds to the climate feedback
    parameter $\lambda$ and the $y$-intercept corresponds to the radiative
    forcing $F_\text{4x}$. These can be used to calculate the \acf{ECS} with
    the Gregory regression method according to \cref{eq:02:ecs}
    \autocite{Gregory2004}. \acs{ECS} is equivalently given by the
    \xintercept{} of the linear regression line divided by $2$. Here,
    $\text{ECS} = 3.14 \unit{K}$.}
  \label{fig:02:gregory_regression}
\end{figure}

For this reason, the Equilibrium Climate Sensitivity is commonly approximated
with the \acf{ECS}, which can be derived from only 150 model years of a
simulation with an abrupt quadrupling of the atmospheric \ac{CO2} concentration
(\nxcotwo{4}) \autocite{Gregory2004}. The basis of the definition of \ac{ECS}
is the simple energy balance model introduced in
\cref{subsec:02:mathematical_framework_feedbacks}. Assuming radiative
equilibrium ($N = 0$), \cref{eq:02:N,eq:02:linear_response} imply
\begin{equation}
  \Delta T = -\frac{F}{\lambda}.
  \label{eq:02:delta_t_eq}
\end{equation}
Thus, the change in \ac{GSAT} in radiative equilibrium can be easily calculated
from the external forcing $F$ and the climate feedback parameter $\lambda$. The
steady state values for $F$ and $\lambda$ can be estimated from a \nxcotwo{4}
simulation that is not in equilibrium by extrapolation with a linear regression
\autocite{Gregory2004}. For this method, the so-called \emph{Gregory
  regression}, the global and annual mean net \ac{TOA} radiation $N$ \vs{} the
change in the annual mean \ac{GSAT} $\Delta T$ for all 150 years of the
\nxcotwo{4} run are plotted (see \cref{fig:02:gregory_regression}). To account
for energy leakage and remove any model drift present in the control climate, a
linear fit of the corresponding pre-industrial control run is subtracted from
the \nxcotwo{4} simulation beforehand \autocite{Andrews2012}. Since the
combination of \cref{eq:02:N,eq:02:linear_response} yields
\begin{equation}
  N = F + \lambda \cdot \Delta T,
  \label{eq:02:N_vs_delta_t}
\end{equation}
$F$ is now given by the \yintercept{} of this linear regression ($F_\text{4x}$)
and $\lambda$ by its slope. Thus, the \ac{ECS} is given by
\begin{equation}
  \text{ECS} = -\frac{F_\text{4x}}{2 \lambda}.
  \label{eq:02:ecs}
\end{equation}
The factor of $2$ in the denominator accounts for the fact the traditional
Equilibrium Climate Sensitivity is defined for an abrupt \ac{CO2} doubling,
whereas here a simulation with an abrupt quadrupling is considered.

@@

\begin{figure}[t]
  \centering
  \includegraphics[width=\FigureWidth{}]{
    ch02_scientific_background/figs/tcr_calculation.pdf}
  \caption{Global and annual mean near-surface air temperature change $\Delta
    T$ for 140 years of a simulation with a $1 \unit{\%}$ per year increase of
    the atmospheric \ac{CO2} concentration (\onepctcotwo{}) for the \acf{CESM}
    of the \acf{NCAR} (blue circles). The temperature change is calculated
    relative to a corresponding pre-industrial control simulation smoothed
    with a linear fit over all 140 years. The \acf{TCR} is defined as the
    temperature change $\Delta T$ at the time of \acs{CO2} doubling averaged
    over a 20 year period (illustrated by the vertical dashed lines). Here,
    $\text{TCR} = 2.35 \unit{K}$ (horizontal dashed line).}
  \label{fig:02:tcr}
\end{figure}

>>>>

\begin{figure}[t]
  \centering
  \includegraphics[width=\FigureWidth{}]{
    ch02_scientific_background/figs/gregory_regression.pdf}
  \caption{Global and annual mean net \acf{TOA} radiation $N$ versus the change
    in global and annual mean near-surface air temperature $\Delta T$ for 150
    years of a simulation with an abrupt quadrupling of the atmospheric
    \acs{CO2} concentration (\nxcotwo{4}) for the \acs{CMIP}5 multi-model mean
    (blue circles). To account for energy leakage and model drift, a linear
    fit of the corresponding pre-industrial control run is subtracted from the
    \nxcotwo{4} simulation. As given by \cref{eq:02:N_vs_delta_t}, the slope
    of the linear regression (black line) corresponds to the climate feedback
    parameter $\lambda$ and the $y$-intercept corresponds to the radiative
    forcing $F_\text{4x}$. These can be used to calculate the \acf{ECS} with
    the Gregory regression method according to \cref{eq:02:ecs}
    \autocite{Gregory2004}. \acs{ECS} is equivalently given by the
    \xintercept{} of the linear regression line divided by $2$. Here,
    $\text{ECS} = 3.14 \unit{K}$.}
  \label{fig:02:gregory_regression}
\end{figure}

For this reason, the Equilibrium Climate Sensitivity is commonly approximated
with the \acf{ECS}, which can be derived from only 150 model years of a
simulation with an abrupt quadrupling of the atmospheric \ac{CO2} concentration
(\nxcotwo{4}) \autocite{Gregory2004}. The basis of the definition of \ac{ECS}
is the simple energy balance model introduced in
\cref{subsec:02:mathematical_framework_feedbacks}. Assuming radiative
equilibrium ($N = 0$), \cref{eq:02:N,eq:02:linear_response} imply
\begin{equation}
  \Delta T = -\frac{F}{\lambda}.
  \label{eq:02:delta_t_eq}
\end{equation}
Thus, the change in \ac{GSAT} in radiative equilibrium can be easily calculated
from the external forcing $F$ and the climate feedback parameter $\lambda$. The
steady state values for $F$ and $\lambda$ can be estimated from a \nxcotwo{4}
simulation that is not in equilibrium by extrapolation with a linear regression
\autocite{Gregory2004}. For this method, the so-called \emph{Gregory
  regression}, the global and annual mean net \ac{TOA} radiation $N$ versus the
change in the annual mean \ac{GSAT} $\Delta T$ for all 150 years of the
\nxcotwo{4} run are plotted (see \cref{fig:02:gregory_regression}). To account
for energy leakage and remove any model drift present in the control climate, a
linear fit of the corresponding pre-industrial control run is subtracted from
the \nxcotwo{4} simulation beforehand \autocite{Andrews2012}. Since the
combination of \cref{eq:02:N,eq:02:linear_response} yields
\begin{equation}
  N = F + \lambda \cdot \Delta T,
  \label{eq:02:N_vs_delta_t}
\end{equation}
$F$ is now given by the \yintercept{} of this linear regression ($F_\text{4x}$)
and $\lambda$ by its slope. Thus, the \ac{ECS} is given by
\begin{equation}
  \text{ECS} = -\frac{F_\text{4x}}{2 \lambda}.
  \label{eq:02:ecs}
\end{equation}
The factor of $2$ in the denominator accounts for the fact the traditional
Equilibrium Climate Sensitivity is defined for an abrupt \ac{CO2} doubling,
whereas here a simulation with an abrupt quadrupling is considered.

@@

\begin{figure}[t]
  \centering
  \includegraphics[width=\FigureWidth{}]{
    ch02_scientific_background/figs/tcr_calculation.pdf}
  \caption{Global and annual mean near-surface air temperature change $\Delta
    T$ for 140 years of a simulation with a $1 \unit{\%}$ per year increase of
    the atmospheric \acs{CO2} concentration (\onepctcotwo{}) for the \acf{CESM}
    of the \acf{NCAR} (blue circles). The temperature change is calculated
    relative to a corresponding pre-industrial control simulation smoothed
    with a linear fit over all 140 years. The \acf{TCR} is defined as the
    temperature change $\Delta T$ at the time of \acs{CO2} doubling averaged
    over a 20 year period (illustrated by the vertical dashed lines). Here,
    $\text{TCR} = 2.35 \unit{K}$ (horizontal dashed line).}
  \label{fig:02:tcr}
\end{figure}

### 2021-01-06 ch03_esmvaltool/esmvaltool.tex
Added figures for ESMValTool contributions (chapter 3)

\begin{figure}[t]
  \centering
  \begin{subfigure}[t]{\SubfigureWidth{}}
    \includegraphics[width=\columnwidth]{ch03_esmvaltool/figs/co2s_map.pdf}
    \caption{}
    \label{fig:03:co2s:a}
  \end{subfigure}
  ~
  \begin{subfigure}[t]{\SubfigureWidth{}}
    \includegraphics[width=\columnwidth]{
      ch03_esmvaltool/figs/co2s_mlo_cycle.pdf}
    \caption{}
    \label{fig:03:co2s:b}
  \end{subfigure}
  \caption{(a) \acs{CMIP}6 \acf{MMM} of the global atmospheric \acs{CO2}
    concentration at surface (\emph{co2s}) averaged over the months \acf{JJA}
    of the year 2014 in the emission-driven historical simulation. The
    \acs{CMIP}6 \ac{MMM} includes the four climate models that provide all
    data needed for the calculation of \emph{co2s} (see legend in (b)). (b)
    Monthly-mean \emph{co2s} at Mauna Loa, Hawaii ($19.5 \unit{\degree N}$,
    $155.6 \unit{\degree W}$; elevation: $3397 \unit{m}$) from 1958 to 2014.
    The thick black line shows observations from the Mauna Loa Observatory
    \autocite{Keeling2005}; the remaining colored lines show emission-driven
    historical simulations from individual \acs{CMIP}6 models.}
  \label{fig:03:co2s}
\end{figure}

@@

>>>>

\begin{figure}[t]
  \centering
  \begin{subfigure}[t]{\SubfigureWidth{}}
    \includegraphics[width=\columnwidth]{ch03_esmvaltool/figs/co2s_map.pdf}
    \caption{}
    \label{fig:03:co2s:a}
  \end{subfigure}
  ~
  \begin{subfigure}[t]{\SubfigureWidth{}}
    \includegraphics[width=\columnwidth]{
      ch03_esmvaltool/figs/co2s_mlo_cycle.pdf}
    \caption{}
    \label{fig:03:co2s:b}
  \end{subfigure}
  \caption{(a) \acs{CMIP}6 \acf{MMM} of the global atmospheric \acs{CO2}
    concentration at surface (\emph{co2s}) averaged over the months \acf{JJA}
    of the year 2014 in the emission-driven historical simulation. The
    \acs{CMIP}6 \acs{MMM} includes the four climate models that provide all
    data needed for the calculation of \emph{co2s} (see legend in (b)). (b)
    Monthly-mean \emph{co2s} at Mauna Loa, Hawaii ($19.5 \unit{\degree N}$,
    $155.6 \unit{\degree W}$; elevation: $3397 \unit{m}$) from 1958 to 2014.
    The thick black line shows observations from the Mauna Loa Observatory
    \autocite{Keeling2005}; the remaining colored lines show emission-driven
    historical simulations from individual \acs{CMIP}6 models.}
  \label{fig:03:co2s}
\end{figure}

@@

\begin{figure}[t]
  \centering
  \begin{subfigure}[t]{\SubfigureWidth{}}
    \includegraphics[width=\columnwidth]{ch03_esmvaltool/figs/cox18_psi.pdf}
    \caption{}
    \label{fig:03:cox18:a}
  \end{subfigure}
  ~
  \begin{subfigure}[t]{\SubfigureWidth{}}
    \includegraphics[width=\columnwidth]{ch03_esmvaltool/figs/cox18_ec.pdf}
    \caption{}
    \label{fig:03:cox18:b}
  \end{subfigure}
  \caption{Emergent constraint on the \acf{ECS} from global temperature
    variability \autocite{Cox2018}. Magenta colors indicate high sensitivity
    climate models ($\lambda < 1 \unit{W m^{-2} K^{-1}}$) and green colors
    indicate low sensitivity climate models ($\lambda > 1 \unit{W m^{-2}
    K^{-1}}$). Both panels were created with the \acs{ESMValTool} and are
    similar to \textcite{Cox2018}. (a) Temperature variability metric $\Psi$
    versus time for historical simulations from \acs{CMIP}5 models (colored
    lines) and observations from the HadCRUT4 dataset (black circles)
    \autocite{Morice2012}. (b) Emergent relationship between \acs{ECS} and the
    temperature variability metric $\Psi$. The black dot-dashed line shows the
    linear regression across the climate model ensemble with its associated
    uncertainties indicated by the black dashed lines (standard prediction
    error; see \cref{eq:02:spe}). Blue lines show the observational constraint
    from HadCRUT4 with its best estimate (dot-dashed line) and standard
    deviation (dashed lines). The letters represent individual \acs{CMIP}5
    models (see \textcite{Cox2018} for details).}
  \label{fig:03:cox18}
\end{figure}

\begin{figure}[t]
  \centering
  \begin{subfigure}[t]{0.34\columnwidth}
    \includegraphics[width=\columnwidth]{ch03_esmvaltool/figs/flato13_942a.pdf}
    \caption{}
    \label{fig:03:flato13:a}
  \end{subfigure}
  ~
  \begin{subfigure}[t]{0.34\columnwidth}
    \includegraphics[width=\columnwidth]{ch03_esmvaltool/figs/flato13_942b.pdf}
    \caption{}
    \label{fig:03:flato13:b}
  \end{subfigure}
  ~
  \begin{subfigure}[t]{0.24\columnwidth}
    \raisebox{7.5mm}{\includegraphics[width=\columnwidth]{
      ch03_esmvaltool/figs/flato13_942_legend.pdf}}
  \end{subfigure}
  \caption{Relationships between climate metrics within the \acs{CMIP}5 climate
    model ensemble. Both panels were created with the \acs{ESMValTool} and are
    similar to \textcite{Flato2013}. (a) \Acf{GSAT} versus the \acf{ECS} for
    \acs{CMIP}5 models for the period \range{1961}{1990} (larger symbols) and
    for pre-industrial control runs (smaller symbols). (b) \acf{TCR} versus
    \acs{ECS} for \acs{CMIP}5 models. The black line shows a linear fit
    between \acs{TCR} and \acs{ECS}.}
\end{figure}
